---
title: "390/Olivia_Handoko"
output: html_document
---
 stem dying/broken below to dbh connection?
- environment?
-quadrant placement?

same tree specie in different cooradinates, how does their dbh differ/if they are alive or if they are dead/other flaw  
- maybe try a couple of different tree species but not all of them 
  - based one species with most stem dead/broken below with those with the least stem dead/broken below in the latest data 
See in each scbi_stem data to see which trees are the least and most for 
- maybe pinpoint a place in the scatterplot or linegraph where the changes are if there is thatmightve affected in the timeline the die off of these trees 
  - consider stem dead and broken below

#notes in code to help choose which trees I guess 
```{r}
#libaries
library(dplyr)
library(tidyverse)
library(tsibble)
library(lubridate)
library(fpp3)
library(fpp2)
library(ggfortify)
library(lubridate)
```

```{r}
#filter out from all scbi_stem data for DF status other than "alive" just to see which trees had the most stem dead and broken below in scbi_stem3 (year 2018)
##

scbi_stem3_notalive <- scbi_stem3 %>%
  filter(DFstatus != "alive") 

scbi_stem3_alive <- scbi_stem3  %>%
  filter(DFstatus != "stem dead" & DFstatus != "broken below" & DFstatus != "prior" & DFstatus != "missing"& DFstatus != "NULL") 

count_stemdead <- scbi_stem3_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead_2018 <- count_stemdead[order(count_stemdead$n),] 

count_brokenbelow <- scbi_stem3_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow_2018 <- count_brokenbelow[order(count_brokenbelow$n),] 

#top 5 from below for broken below and stem dead (some tree species cant be found in broken below so use top 10 that is in both data)
#libel: stem dead - 11084, broken below: 1025 (the most for both)
#astr: stem dead 1988 broken below: 110 (the 2nd for both)
#caca: stem dead 747 broken below 35 (3rd for both)
#unk: stem dead 677 broken below 1 (4th for stem dead but the least for broken below)
#4th for broken below is: ceca at 35 and stem dead at 529 (7th for stem dead)
#havi: stem dead at 570 broken below at 20 (5th for stem dead but 7th for broken below)
#5th for broken below is litu at 27 and 267 at stem dead (10th for stem dead)

# #Stem dead 
# 5 havi	stem dead	570 -> 7th most for broken below 
# 4 unk	stem dead	677 -> least for broken below
# 3 caca	stem dead	747
# 2 astr	stem dead	1988
# 1 libe stem dead	11084
#______________________________________
#(for reference to previous data - 2013)
# cofl stem dead: 501 8th most for 2018, 
#7th is ceca at 529 in 2018
#ulru stem dead: 544 6th most for 2018
#_____________________________________
#reference to data - 2008)
#stem dead ceca 529 at 7th place in 2018
#vipr stem dead 157 at 15th place in 2018
#________________________________________
# #broken below 
# 5 litu	broken below	27 -> 10th most for stem dead 
# 4 ceca	broken below	31 -> 7th most for stem dead 
# 3 caca	broken below	35
# 2 astr	broken below	110
# 1 libe	broken below	1025

```

```{r}
#do same as above but for scbi_stem data 2 (year 2013)
#filtered out  df status prior and missing
scbi_stem2_notalive <- scbi_stem2 %>%
  filter(DFstatus != "alive" & DFstatus != "prior" & DFstatus != "missing" & DFstatus != "NULL") 

scbi_stem2_alive <- scbi_stem2  %>%
  filter(DFstatus != "stem dead" & DFstatus != "broken below" & DFstatus != "prior" & DFstatus != "missing") 

count_stemdead_2 <- scbi_stem2_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead2_2013 <- count_stemdead_2[order(-count_stemdead_2$n),] 

count_brokenbelow_2 <- scbi_stem2_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow2_2013 <- count_brokenbelow_2[order(-count_brokenbelow_2$n),] 

#Stem dead 
# 1 libe	stem dead	3660
# 2	astr	stem dead	869
# 3	unk	stem dead	676
# 4	ulru	stem dead	433
# 5	cofl	stem dead	390
#______________________________________
#(reference for 2018)
#havi stem dead 318, 6th most in 2013
#ceca stem dead at 233, 8th most in 2013
#7th is caca in stem dead at 304 in 2013
#______________________________________

#(reference for 2008)
#Vipr stem dead 91 at 16th place 2013
#Ceca stem dead 233, 8th place 2013
#Havi stem dead 318 at 6th place 2013

#______________________________________
#broken below 
# 1 libe	broken below	694
# 2	astr	broken below	107
# 3	ceca	broken below	91
# 4	caca	broken below	46
# 5	litu	broken below	39
#_____________________________________
#(referencing for 2008)
#havi broken below 36 ranked 7th for 2013
```


```{r}
#do same as above but for scbi_stem data 1 (year 2008)
#filtered out  df status prior and missing
scbi_stem1_notalive <- scbi_stem1 %>%
  filter(DFstatus != "alive" & DFstatus != "prior" & DFstatus != "missing" & DFstatus != "NULL") 

scbi_stem1_alive <- scbi_stem1  %>%
  filter(DFstatus != "stem dead" & DFstatus != "broken below" & DFstatus != "prior" & DFstatus != "missing") 

count_stemdead_1 <- scbi_stem1_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead1_2008 <- count_stemdead_1[order(-count_stemdead_1$n),] 

count_brokenbelow_1 <- scbi_stem1_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow1_2008 <- count_brokenbelow_1[order(-count_brokenbelow_1$n),] 

#Stem dead 
# 1 libe	stem dead	156
# 2 havi	stem dead	36
# 3 ceca	stem dead	19
# 4 vipr	stem dead	18
# 5	astr	stem dead	9
#_________________________
#(reference for 2018)
#caca stem dead 7, tied at 7th place in 2008
#unk stem dead 6, tied at 10th in 2008
#_________________________
#(reference for 2013)
#unk stem dead 6, tied at 10th in 2008
#ulru stem dead 2, tied at 11th? in 2008
#cofl stem dead 6, tied at 7th? in 2008
#_________________________
#Broken below 
# 1 havi	broken below	24
# 2	libe	broken below	19
# 3	ceca	broken below	17
# 4	astr	broken below	11
# 5	caca	broken below	5
#_________________________
#(reference for 2018)
#litu broken below 1 ranked 7th but there are about 13 with just 1 in 2008
```

```{r}
##trees we will be using bc of stem dead and broken below
# Unk	
# Caca	
# Astr
# Libe 
# Litu	 
# Ulru	
# Cofl	
# Vipr 
# Ceca 
# Havi 
```
______________________________________________________________________________________
```{r}
#thought id filter for just the above trees but i guess it's good to leave the rest of the tree species here to see how they hold up in comparison with them too

new_scbi_stem1_notalive <- scbi_stem1_notalive %>% 
    filter(sp %in% c("libe")) 
#  filter(StemTag %in% c(1,2)) %>%
  

new_scbi_stem1_notalive <-new_scbi_stem1_notalive[duplicated(new_scbi_stem1_notalive$tag), ]
new_scbi_stem1_notalive <-new_scbi_stem1_notalive[duplicated(new_scbi_stem1_notalive$treeID), ]

new_scbi_stem1_notalive <- new_scbi_stem1_notalive[order(as.Date(new_scbi_stem1_notalive$ExactDate, format="%m/%d/%Y")),]


#summarise(group_by(new_scbi_stem1_notalive,sp),length(sp))

new_scbi_stem1_alive <-scbi_stem1_alive %>%
  filter(sp %in% c("libe")) %>%
#  filter(StemTag %in% c(1,2)) %>%
  filter(quadrat <=300 | quadrat >= 1000 ) %>%
  filter(treeID <= 300)
new_scbi_stem1_alive <-new_scbi_stem1_alive[duplicated(new_scbi_stem1_alive$tag), ]
new_scbi_stem1_alive <-new_scbi_stem1_alive[duplicated(new_scbi_stem1_alive$treeID), ]

new_scbi_stem1_alive <- new_scbi_stem1_alive[order(as.Date(new_scbi_stem1_alive$ExactDate, format="%m/%d/%Y")),]

#summarise(group_by(new_scbi_stem1_alive,sp),length(sp))

new_scbi_stem2_notalive <- scbi_stem2_notalive %>% 
    filter(sp %in% c("libe")) %>%
    filter (!dbh == 0)
 #   filter(StemTag %in% c(1,2)) %>%



new_scbi_stem2_alive <-scbi_stem2_alive %>%
  filter(sp %in% c( "libe")) %>%
 # filter(StemTag %in% c(1,2)) %>%
  filter(quadrat <=300 | quadrat >= 1000) %>%
  filter(treeID <= 300)


new_scbi_stem2_alive <-new_scbi_stem2_alive[duplicated(new_scbi_stem2_alive$tag), ]
new_scbi_stem2_alive <-new_scbi_stem2_alive[duplicated(new_scbi_stem2_alive$treeID), ]

new_scbi_stem2_alive <- new_scbi_stem2_alive[order(as.Date(new_scbi_stem2_alive$ExactDate, format="%m/%d/%Y")),]

new_scbi_stem3_notalive <- scbi_stem3_notalive %>% 
    filter(sp %in% c("libe")) %>%
  filter(DFstatus != "alive" & DFstatus != "prior" & DFstatus != "missing" & dbh != "NULL") 

 # filter(StemTag %in% c(1,2)) %>%


new_scbi_stem3_alive <-scbi_stem3_alive %>%
  filter(sp %in% c( "libe")) %>%
  filter(StemTag %in% c(1,2)) %>%
  filter(quadrat <=2000)
#new_scbi_stem3_alive <-new_scbi_stem3_alive[!duplicated(new_scbi_stem3_alive$tag), ]
```

```{r}
#bind the alive and not alive data together 
##
total_alive <- rbind(new_scbi_stem1_alive, new_scbi_stem2_alive) %>% select("sp", "quadrat", "dbh", "ExactDate","date","quadrat", "treeID", "StemTag", "tag", "MeasureID") %>%
  filter(!dbh %in% c(0, "NULL"))  %>%
  filter(StemTag != 10) %>%
  group_by(treeID, StemTag) %>%
  filter(n() >= 2 ) 
#I want tree ID and StemTag frequency to be more than just one for ST and 3 for treeID

   #filter(!treeID %in% c(3,13,14,15,18,35,40,42,53,56,58,67,68,87,105,110,189,191,200,201,209,213,214,219,222,230,234,236,251,253,258,259,260,262,269,272,274,276,280,281,282,287,288,289,290,297))

as.data.frame((table(total_alive$treeID))) 
as.data.frame((table(total_alive$StemTag)))

total_alive %>% group_by(treeID, StemTag) %>% summarize(freq = n()) %>%
  filter(freq >= 2)

# total_alive <-total_alive[duplicated(treeID$StemTag), ] %>%
#   filter(tag != 20001)

total_notalive <- rbind(new_scbi_stem2_notalive, new_scbi_stem1_notalive, new_scbi_stem3_notalive) %>% select("sp", "dbh", "ExactDate", "quadrat", "date", "treeID", "StemTag", "tag", "MeasureID") %>%
  filter(!dbh %in% c(0, "NULL")) 

```

```{r}
ggplot(total_alive) + 
  geom_line(aes(x=date, y=dbh, group = interaction(treeID, StemTag, quadrat), color = quadrat)) + facet_wrap(~quadrat)



#+
  #geom_text(data = subset(new_scbi_stem1_notalive, ExactDate== "12/18/2008"), aes(label = quadrat, x = ExactDate, y = dbh))


```
```{r}
ggplot(new_scbi_stem1_notalive) + 
  geom_point(aes(x=date, y=dbh, group = treeID, color = quadrat))
```
```{r}
ggplot(new_scbi_stem2_notalive) + 
  geom_point(aes(x=date, y=dbh, group = treeID, color = quadrat))
```


```{r}
ggplot(new_scbi_stem3_notalive) + 
  geom_point(aes(x=date, y=dbh, group = treeID, color = quadrat)) 
```

Honestly, this looked way better in my head. I chose to just focus on libe and see if libes placed in different quadrants would impact its growth or if they were more likely to die there (and when), considering tree ID and Stem Tag so that I can differentiate them from like a libe with the same StemTag but different TreeID to match with data from the later years with HOPEFULLY the same tree.  I chose qudrants from the lowest to the highest and not considering anything in between because the lines were getting too intense for the alive libe trees and I thought it'd be a good contrast to see clearly.
Maybe from this we can find average dbh of libes dead and in which quadrat that this mostly likely happens in. Maybe you can also find how the seasons affect this too. 
Like maybe libe in a certain quadrat do well in one season but not so much in another

The above r code, I didn't really use all of them I think. I had an idea and then I started running them then realized problems and then my idea completely changed 360 problem after problem and then I ended up here. 

Also, weird. I'm looking at the first graph and some of the dbh went down. In what case would that happen? Like a tree can grow smaller? I also checked to see if it was a r code mistake but I'm looking at treeID 263 and StemTag 9 and it does go down over the years by .6