---
title: "390/Olivia_Handoko"
output: html_document
---
 stem dying/broken below to dbh connection?
- environment?
-quadrant placement?

same tree specie in different cooradinates, how does their dbh differ/if they are alive or if they are dead/other flaw  
- maybe try a couple of different tree species but not all of them 
  - based one species with most stem dead/broken below with those with the least stem dead/broken below in the latest data 
See in each scbi_stem data to see which trees are the least and most for 
- maybe pinpoint a place in the scatterplot or linegraph where the changes are if there is thatmightve affected in the timeline the die off of these trees 
  - consider stem dead and broken below

#notes in code to help choose which trees I guess 
```{r}
#libaries
library(dplyr)

```

```{r}
#filter out from all scbi_stem data for DF status other than "alive" just to see which trees had the most stem dead and broken below in scbi_stem3 (year 2018)
##

scbi_stem3_notalive <- scbi_stem3 %>%
  filter(DFstatus != "alive") 

count_stemdead <- scbi_stem3_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead_2018 <- count_stemdead[order(count_stemdead$n),] 

count_brokenbelow <- scbi_stem3_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow_2018 <- count_brokenbelow[order(count_brokenbelow$n),] 

#top 5 from below for broken below and stem dead (some tree species cant be found in broken below so use top 10 that is in both data)
#libel: stem dead - 11084, broken below: 1025 (the most for both)
#astr: stem dead 1988 broken below: 110 (the 2nd for both)
#caca: stem dead 747 broken below 35 (3rd for both)
#unk: stem dead 677 broken below 1 (4th for stem dead but the least for broken below)
#4th for broken below is: ceca at 35 and stem dead at 529 (7th for stem dead)
#havi: stem dead at 570 broken below at 20 (5th for stem dead but 7th for broken below)
#5th for broken below is litu at 27 and 267 at stem dead (10th for stem dead)

# #Stem dead 
# 5 havi	stem dead	570 -> 7th most for broken below 
# 4 unk	stem dead	677 -> least for broken below
# 3 caca	stem dead	747
# 2 astr	stem dead	1988
# 1 libe stem dead	11084
#______________________________________
#(for reference to previous data - 2013)
# cofl stem dead: 501 8th most for 2018, 
#7th is ceca at 529 in 2018
#ulru stem dead: 544 6th most for 2018
#_____________________________________
#reference to data - 2008)
#stem dead ceca 529 at 7th place in 2018
#vipr stem dead 157 at 15th place in 2018
#________________________________________
# #broken below 
# 5 litu	broken below	27 -> 10th most for stem dead 
# 4 ceca	broken below	31 -> 7th most for stem dead 
# 3 caca	broken below	35
# 2 astr	broken below	110
# 1 libe	broken below	1025

```

```{r}
#do same as above but for scbi_stem data 2 (year 2013)
#filtered out  df status prior and missing
scbi_stem2_notalive <- scbi_stem2 %>%
  filter(DFstatus != "alive" & DFstatus != "prior" & DFstatus != "missing") 

count_stemdead_2 <- scbi_stem2_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead2_2013 <- count_stemdead_2[order(-count_stemdead_2$n),] 

count_brokenbelow_2 <- scbi_stem2_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow2_2013 <- count_brokenbelow_2[order(-count_brokenbelow_2$n),] 

#Stem dead 
# 1 libe	stem dead	3660
# 2	astr	stem dead	869
# 3	unk	stem dead	676
# 4	ulru	stem dead	433
# 5	cofl	stem dead	390
#______________________________________
#(reference for 2018)
#havi stem dead 318, 6th most in 2013
#ceca stem dead at 233, 8th most in 2013
#7th is caca in stem dead at 304 in 2013
#______________________________________

#(reference for 2008)
#Vipr stem dead 91 at 16th place 2013
#Ceca stem dead 233, 8th place 2013
#Havi stem dead 318 at 6th place 2013

#______________________________________
#broken below 
# 1 libe	broken below	694
# 2	astr	broken below	107
# 3	ceca	broken below	91
# 4	caca	broken below	46
# 5	litu	broken below	39
#_____________________________________
#(referencing for 2008)
#havi broken below 36 ranked 7th for 2013
```


```{r}
#do same as above but for scbi_stem data 1 (year 2008)
#filtered out  df status prior and missing
scbi_stem1_notalive <- scbi_stem1 %>%
  filter(DFstatus != "alive" & DFstatus != "prior" & DFstatus != "missing") 

count_stemdead_1 <- scbi_stem1_notalive %>%
  filter(DFstatus == "stem dead") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_stemdead1_2008 <- count_stemdead_1[order(-count_stemdead_1$n),] 

count_brokenbelow_1 <- scbi_stem1_notalive %>%
  filter(DFstatus == "broken below") %>%
  group_by(sp) %>%
  count(DFstatus) 
count_brokenbelow1_2008 <- count_brokenbelow_1[order(-count_brokenbelow_1$n),] 

#Stem dead 
# 1 libe	stem dead	156
# 2 havi	stem dead	36
# 3 ceca	stem dead	19
# 4 vipr	stem dead	18
# 5	astr	stem dead	9
#_________________________
#(reference for 2018)
#caca stem dead 7, tied at 7th place in 2008
#unk stem dead 6, tied at 10th in 2008
#_________________________
#(reference for 2013)
#unk stem dead 6, tied at 10th in 2008
#ulru stem dead 2, tied at 11th? in 2008
#cofl stem dead 6, tied at 7th? in 2008
#_________________________
#Broken below 
# 1 havi	broken below	24
# 2	libe	broken below	19
# 3	ceca	broken below	17
# 4	astr	broken below	11
# 5	caca	broken below	5
#_________________________
#(reference for 2018)
#litu broken below 1 ranked 7th but there are about 13 with just 1 in 2008
```

```{r}
##trees we will be using bc of stem dead and broken below
# Unk	
# Caca	
# Astr
# Libe 
# Litu	 
# Ulru	
# Cofl	
# Vipr 
# Ceca 
# Havi 
```
______________________________________________________________________________________
```{r}

```

